---
title: "menarche_research"
output: html_document
date: "2022-10-10"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(table1)
library(MatchIt)
library(skimr)
library(broom)
library(gtsummary)
library(aod)
```


```{r, echo = FALSE}
menarche_data <- read.delim('/Users/spencerfriedman/Documents/School/Spracklen_lab/f31_ctos_inv.dat')
demo_data <- read.delim('/Users/spencerfriedman/Documents/School/Spracklen_lab/dem_ctos_inv.dat')
socio_data <- read.delim('/Users/spencerfriedman/Documents/School/Spracklen_lab/f34_ctos_inv.dat')
race_data <- read.delim('/Users/spencerfriedman/Documents/School/Spracklen_lab/f41_imputed_ctos_inv.dat')
preg_data <- read.delim('/Users/spencerfriedman/Documents/School/Spracklen_lab/f158_ctos_inv.dat')
measure_data <- read.delim('/Users/spencerfriedman/Documents/School/Spracklen_lab/f80_ctos_inv.dat')
nses_data <- read.csv('/Users/spencerfriedman/Documents/School/Spracklen_lab/ct_os_tiv.csv')
```

```{r}
glimpse(nses_data)
glimpse(menarche_data)
```

```{r}
measure_data <- filter(measure_data, F80VTYP == 1)
menarche_data <- menarche_data %>% drop_na(MENARCHE)
names(nses_data)[names(nses_data) == 'id'] <- 'ID'
```


glimpse(measure_data)



```{r}
merged <- left_join(x = preg_data, y = race_data, by='ID')

merged <- left_join(x = merged, y = socio_data, by ="ID" )

merged <- left_join(x = merged, y = menarche_data, by ="ID")

merged <- left_join(x = merged, y = demo_data, by = "ID")

merged <- left_join(x = merged, y = measure_data, by='ID')

merged <- left_join(x = merged, y = nses_data, by = 'ID')
```

merged <- left_join(x = merged, y = socio_data, by = "ID")

```{r}
glimpse(merged)
```


skimr::skim(merged)
names(merged)


```{r}

merged$MENARCHE <- merged$MENARCHE + 8
fortify(merged)

```

```{r}
ggplot(merged, aes(x = MENARCHE)) + geom_histogram(bins = 9)
```

```{r}
se <- function(x) sqrt(var(x)/length(x))
```



```{r}

```

```{r}
merged$MENARCHE <- ifelse(merged$MENARCHE <= 11, '<=11', 
                   ifelse(merged$MENARCHE == 12, '12', 
                   ifelse(merged$MENARCHE == 13, '13', 
                   ifelse(merged$MENARCHE >= 14,'>=14', 0))))

merged$ethnicity <- ifelse(merged$ETHNICITY_IMPUTED == 0, 0, 1)

merged$ALCOHOL <- ifelse(merged$ALCOHOL == 1,1, 
                         ifelse(merged$ALCOHOL ==2, 2, 3))

merged$RACENIH <- ifelse(merged$RACENIH == 2, 1,
                         ifelse(merged$RACENIH == 3, 1,
                                ifelse(merged$RACENIH == 4, 2, 
                                       ifelse(merged$RACENIH == 5, 3, 4))))

merged$EDUC <- ifelse(merged$EDUC <= 5, 1, 
                      ifelse(merged$EDUC == 6, 2,
                             ifelse(merged$EDUC == 7, 2,
                             ifelse(merged$EDUC >= 8, 3, 0))))

ggplot(merged, aes(x = MENARCHE)) + geom_histogram(stat= 'count')
```
```{r}
merged$EDUC <- factor(merged$EDUC, levels = c(1,2,3), labels = c('<= High school diploma/GED','School after high school','College degree or higher'))
merged$MENARCHE <- factor(merged$MENARCHE, levels = c('<=11', '12','13','>=14'), labels = c('<=11', '12','13','>=14'))

merged$PREECLMP <- factor(merged$PREECLMP, levels = c(0,1), labels = c('No Preeclampsia during pregnancy','Preeclampsia during pregnancy'))

merged$HYPTPREG <- factor(merged$HYPTPREG, levels= c(0,1), labels = c('No Hypertension during preg','Hypertension during preg'))

merged$DIABPREG <- factor(merged$DIABPREG, levels= c(0,1), labels = c('No Diabetes during preg','Diabetes during preg'))

merged$BABYGT9LBS <- factor(merged$BABYGT9LBS, levels = c(0,1), labels = c("No Overweight Baby", "Overweight baby") )

merged$BABYLT5LBS <- factor(merged$BABYLT5LBS, levels = c(0,1), labels = c("No Underweight baby",'Underweight baby'))

merged$ALCOHOL <- factor(merged$ALCOHOL, levels = c(1,2,3), labels = c('Non drinker','Past drinker','Current drinker'))

merged$SMOKING <- factor(merged$SMOKING, levels = c(0,1,2), labels = c('Non smoker','Past smoker','Current smoker'))

merged$ETHNICITY <- factor(merged$ETHNICNIH, levels = c(0,1), labels = c('Not Hispanic/Latino','Hispanic/Latino'))
  
merged$RACE <- factor(merged$RACENIH, levels = c(1,2,3,4), labels = c('Asian/Pacific Islander','Black','White','Other/Unknown'))

merged$PRETERM <- ifelse(merged$PREGNUM - merged$PREG6MN > 0, 1, 0)

merged$PRETERM <- factor(merged$PRETERM, levels = c(0,1), labels = c('No preterm birth
','Preterm birth'))

label(merged$t0.normindx) <- 'Normalized Socioeconomic Status (NSES)'

```

```{r}
my.render.cont <- function(x) {
    with(stats.apply.rounding(stats.default(x), digits=2), c("",
        "Mean (SD)"=sprintf("%s (&plusmn; %s)", MEAN, SD)))
}
my.render.cat <- function(x) {
    c("", sapply(stats.default(x), function(y) with(y,
        sprintf("%d (%0.0f %%)", FREQ, PCT))))
}

pvalue <- function(x, ...) {
    # Construct vectors of data y, and groups (strata) g
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        # For numeric variables, perform a standard 2-sample t-test
        p <- t.test(y ~ g)$p.value
    } else {
        # For categorical variables, perform a chi-squared test of independence
        p <- chisq.test(table(y, g))$p.value
    }
    # Format the p-value, using an HTML entity for the less-than sign.
    # The initial empty string places the output on the line below the variable label.
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}

```

```{r}
table1(~ AGE + MENARCHE + ETHNICITY + EDUC +  RACE + BMI + t0.normindx + SMOKING + ALCOHOL| PREECLMP, data = merged, overall= F, extra.col=list('P-value'=pvalue), render.continuous = my.render.cont, render.categorical = my.render.cat)

table1(~ AGE + MENARCHE + ETHNICITY + EDUC + RACE + BMI + t0.normindx + SMOKING + ALCOHOL| HYPTPREG, data = merged,render.continuous = my.render.cont, render.categorical = my.render.cat, overall= F, extra.col=list('P-value'=pvalue))

table1(~ AGE + MENARCHE + ETHNICITY + EDUC + RACE + BMI + t0.normindx + SMOKING + ALCOHOL| BABYLT5LBS, data = merged,render.continuous = my.render.cont, render.categorical = my.render.cat, overall= F, extra.col=list('P-value'=pvalue))

table1(~ AGE + MENARCHE + ETHNICITY + EDUC + RACE + BMI + t0.normindx + SMOKING + ALCOHOL| BABYGT9LBS, data = merged,render.continuous = my.render.cont, render.categorical = my.render.cat, overall= F, extra.col=list('P-value'=pvalue))

table1(~ AGE + MENARCHE + ETHNICITY + EDUC + RACE + BMI + t0.normindx + SMOKING + ALCOHOL| DIABPREG, data = merged,render.continuous = my.render.cont, render.categorical = my.render.cat, overall= F, extra.col=list('P-value'=pvalue))

table1(~ AGE + MENARCHE + ETHNICITY + EDUC + RACE + BMI + t0.normindx + SMOKING + ALCOHOL| PRETERM, data = merged,render.continuous = my.render.cont, render.categorical = my.render.cat, overall= F, extra.col=list('P-value'=pvalue))
```


univ_tab <- linelist %>% 
  dplyr::select(MENARCHE, PREECLMP) %>% ## select variables of interest

  tbl_uvregression(                         ## produce univariate table
    method = glm,                           ## define regression want to run (generalised linear model)
    y = PREECLMP,                            ## define outcome variable
    method.args = list(family = binomial),  ## define what type of glm want to run (logistic)
    exponentiate = TRUE                     ## exponentiate to produce odds ratios (rather than log odds)
  )

## view univariate results table 
univ_tab


```{r undajusted regression}
merged$MENARCHE <- relevel(merged$MENARCHE, ref='12')

modelPrec <- glm(PREECLMP ~ MENARCHE, family = 'binomial', data = merged)
exp(cbind(OR = coef(modelPrec), confint(modelPrec)))
modelPrec %>% tbl_regression(pvalue_fun = ~style_pvalue(.x, digits = 3), exponentiate = TRUE) %>% add_global_p() 
summary(modelPrec)

modelHypr <- glm(HYPTPREG ~ MENARCHE, family = 'binomial', data = merged)
exp(cbind(OR = coef(modelHypr), confint(modelHypr)))
modelHypr %>% tbl_regression(pvalue_fun = ~style_pvalue(.x, digits = 3), exponentiate = TRUE) %>% add_global_p() 
summary(modelHypr)

modelUnder <- glm(BABYLT5LBS ~ MENARCHE, family = 'binomial', data = merged)
exp(cbind(OR = coef(modelUnder), confint(modelUnder)))
modelUnder %>% tbl_regression(pvalue_fun = ~style_pvalue(.x, digits = 3), exponentiate = TRUE) %>% add_global_p() 
summary(modelUnder)

modelOver <- glm(BABYGT9LBS ~ MENARCHE, family = 'binomial', data = merged)
exp(cbind(OR = coef(modelOver), confint(modelOver)))
modelOver %>% tbl_regression(pvalue_fun = ~style_pvalue(.x, digits = 3), exponentiate = TRUE) %>% add_global_p() 
summary(modelOver)
modelDiab <- glm(DIABPREG ~ MENARCHE, family = 'binomial', data = merged)
exp(cbind(OR = coef(modelDiab), confint(modelDiab)))
modelDiab %>% tbl_regression(pvalue_fun = ~style_pvalue(.x, digits = 3), exponentiate = TRUE) %>% add_global_p() 
summary(modelDiab)

modelPret <- glm(PRETERM ~ MENARCHE, family = 'binomial', data = merged)
exp(cbind(OR= coef(modelPret), confint(modelPret)))
modelPret %>% tbl_regression(pvalue_fun = ~style_pvalue(.x, digits = 3), exponentiate = TRUE) %>% add_global_p() 
summary(modelPret)
```

```{r demo adjusted regression}
modelPrec <- glm(PREECLMP ~ MENARCHE + AGE + RACE + ETHNICITY + t0.normindx + EDUC, family = 'binomial', data = merged)
exp(cbind(OR = coef(modelPrec), confint(modelPrec)))
modelPrec %>% tbl_regression(pvalue_fun = ~style_pvalue(.x, digits = 3), exponentiate = TRUE) %>% add_global_p() 
summary(modelPrec)
```
```{r}
modelHypr <- glm(HYPTPREG ~ MENARCHE + AGE + RACE + ETHNICITY + t0.normindx + EDUC, family = 'binomial', data = merged)
exp(cbind(OR = coef(modelHypr), confint(modelHypr)))
modelHypr %>% tbl_regression(pvalue_fun = ~style_pvalue(.x, digits = 3), exponentiate = TRUE) %>% add_global_p() 
summary(modelHypr)
```
```{r}
modelUnder <- glm(BABYLT5LBS ~ MENARCHE + AGE + RACE + ETHNICITY + t0.normindx + EDUC, family = 'binomial', data = merged)
exp(cbind(OR = coef(modelUnder), confint(modelUnder)))
modelUnder %>% tbl_regression(pvalue_fun = ~style_pvalue(.x, digits = 3), exponentiate = TRUE) %>% add_global_p() 
summary(modelUnder)
```
```{r}
modelOver <- glm(BABYGT9LBS ~ MENARCHE + AGE + RACE + ETHNICITY + t0.normindx + EDUC, family = 'binomial', data = merged)
exp(cbind(OR = coef(modelOver), confint(modelOver)))
modelOver %>% tbl_regression(pvalue_fun = ~style_pvalue(.x, digits = 3), exponentiate = TRUE) %>% add_global_p() 
summary(modelOver)
```
```{r}
modelDiab <- glm(DIABPREG ~ MENARCHE + AGE + RACE + ETHNICITY + t0.normindx + EDUC, family = 'binomial', data = merged)
exp(cbind(OR = coef(modelDiab), confint(modelDiab)))
modelDiab %>% tbl_regression(pvalue_fun = ~style_pvalue(.x, digits = 3), exponentiate = TRUE) %>% add_global_p() 
summary(modelDiab)
```
```{r}
modelPret <- glm(PRETERM ~ MENARCHE + AGE + RACE + ETHNICITY + t0.normindx + EDUC, family = 'binomial', data = merged)
exp(cbind(OR= coef(modelPret), confint(modelPret)))
modelPret %>% tbl_regression(pvalue_fun = ~style_pvalue(.x, digits = 3), exponentiate = TRUE) %>% add_global_p() 
summary(modelPret)
```

```{r demo and life edjusted}
modelPrec <- glm(PREECLMP ~ MENARCHE + AGE + RACE + ETHNICITY + t0.normindx + EDUC + BMI + SMOKING + ALCOHOL, family = 'binomial', data = merged)
exp(cbind(OR = coef(modelPrec), confint(modelPrec)))
modelPrec %>% tbl_regression(pvalue_fun = ~style_pvalue(.x, digits = 3), exponentiate = TRUE) %>% add_global_p() 
summary(modelPrec)
```
```{r}
modelHypr <- glm(HYPTPREG ~ MENARCHE + AGE + RACE + ETHNICITY + t0.normindx + EDUC + BMI + SMOKING + ALCOHOL, family = 'binomial', data = merged)
exp(cbind(OR = coef(modelHypr), confint(modelHypr)))
modelHypr %>% tbl_regression(pvalue_fun = ~style_pvalue(.x, digits = 3), exponentiate = TRUE) %>% add_global_p() 
summary(modelHypr)
```
```{r}
modelUnder <- glm(BABYLT5LBS ~ MENARCHE + AGE + RACE + ETHNICITY + t0.normindx + EDUC + BMI + SMOKING + ALCOHOL, family = 'binomial', data = merged)
exp(cbind(OR = coef(modelUnder), confint(modelUnder)))
modelUnder %>% tbl_regression(pvalue_fun = ~style_pvalue(.x, digits = 3), exponentiate = TRUE) %>% add_global_p() 
summary(modelUnder)
```
```{r}
modelOver <- glm(BABYGT9LBS ~ MENARCHE + AGE + RACE + ETHNICITY + t0.normindx + EDUC + BMI + SMOKING + ALCOHOL, family = 'binomial', data = merged)
exp(cbind(OR = coef(modelOver), confint(modelOver)))
modelOver %>% tbl_regression(pvalue_fun = ~style_pvalue(.x, digits = 3), exponentiate = TRUE) %>% add_global_p() 
summary(modelOver)
```
```{r}
modelDiab <- glm(DIABPREG ~ MENARCHE + AGE + RACE + ETHNICITY + t0.normindx + EDUC + BMI + SMOKING + ALCOHOL, family = 'binomial', data = merged)
exp(cbind(OR = coef(modelDiab), confint(modelDiab)))
modelDiab%>% tbl_regression(pvalue_fun = ~style_pvalue(.x, digits = 3), exponentiate = TRUE) %>% add_global_p() 
summary(modelDiab)
```
```{r}
modelPret <- glm(PRETERM ~ MENARCHE + AGE + RACE + ETHNICITY + t0.normindx + EDUC + BMI + SMOKING + ALCOHOL, family = 'binomial', data = merged)
exp(cbind(OR= coef(modelPret), confint(modelPret)))
modelPret %>% tbl_regression(pvalue_fun = ~style_pvalue(.x, digits = 3), exponentiate = TRUE) %>% add_global_p() 
summary(modelPret)

```